### 2019-07-23 paper/paper.tex
ADD: initial draft of introduction

\documentclass[conference,compsoc,a4paper]{IEEEtran}
\usepackage[utf8]{inputenc}

% configs from ba
\usepackage[english]{babel}
\usepackage[raiselinks=true,
			bookmarks=true,
			bookmarksopenlevel=1,
			bookmarksopen=true,
			bookmarksnumbered=true,
			hyperindex=true,
			plainpages=false,
			pdfpagelabels=true,
			pdfborder={0 0 0.5},
			colorlinks=false,
			linkbordercolor={0 0.61 0.50},
			citebordercolor={0 0.61 0.50}]{hyperref}
\addto\extrasenglish{%
  \def\chapterautorefname{Chapter}%
  \def\chapterrefname{Chapter}%
  \def\sectionautorefname{Section}%
  \def\sectionrefname{Section}%
  \def\subsectionautorefname{Section}%
  \def\subsectionrefname{Section}%
  \def\subsubsectionautorefname{Section}%
  \def\subsubsectionrefname{Section}%
  \def\paragraphautorefname{Paragraph}%
  \def\paragraphrefname{Paragraph}%
}

\ifCLASSOPTIONcompsoc
  % IEEE Computer Society needs nocompress option
  % requires cite.sty v4.0 or later (November 2003)
  \usepackage[nocompress]{cite}
\else
  % normal IEEE
  \usepackage{cite}
\fi

% correct bad hyphenation here
\hyphenation{op-tical net-works semi-conduc-tor}

\begin{document}

\title{Data-parallel Training of Neural Networks}

% author names and affiliations
% use a multiple column layout for up to three different
% affiliations
\author{\IEEEauthorblockN{Katharina Flügel}
\IEEEauthorblockA{TODO?}}

% make the title area
\maketitle

% General rule: no math, special symbols or citations in the abstract
\begin{abstract}
TODO Abstract
\end{abstract}

\section{Introduction} % (fold)
\label{sec:introduction}
% Section Road Map:
% Motivation
%   1. Neural networks and deep learning are important and have many applications.
%   2. Training neural networks is time consuming due to evergrowing models (parameters) and datasets
%      (number and size of inputs (e.g. images)), and there is reason to speed this up.
%   3. Parallelizing and distributing the training is a/the solution to this.
%
% Contributions
%   1. Comparision of multiple parallelism approaches for distributed deep learning
%   2. Implementation of one such technique (synchronous SGD with all-reduce). Evaluation on image classicication datasets.
%   3. Results: TODO: main findings
%
% Paper Road Map
% 	1. Short road map containing the remaining sections

%% Motivation
Neural networks have become an essential factor in solving many practical challenges.
There is a multitude of applications from computer vision, audio and speech processing, to processing text and natural language \cite{chen2016-Revisiting-distributed-synchronous-SGD,dean2012-Large-scale-distributed,yadan2013-Multi-GPU-Training}.
More practical applications include machine translation \cite{sutskever2014sequence}, self-driving cars, and automatic driver assistance systems \cite{jin2016-How-to-scale}.
%
In this paper, we focus mostly on the convolutional neural networks (CNNs) used in computer vision.
Typical tasks include image classification, character and object recognition, and object detection \cite{paine2013-GPU-async-SGD}.

The recent advances in deep learning are facilitated by a significant increase in both model size and available training data, accompanied further by the increased computing power \cite{chen2016-Revisiting-distributed-synchronous-SGD,chilimbi2014-Project-Adam}.
%
While datasets used to be rather small contain tens of thousands of images \cite{lecun1998gradient,krizhevsky2009CIFAR10}, recent datasets like ImageNet \cite{imagenet_cvpr09} contain millions of images for thousands of classes.
Moreover, the internal datasets of some companies are said to contain billions of images with hundreds of thousands of classes \cite{iandola2016-Firecaffe}.
%
To accommodate these massive datasets models with larger learning capacities are required \cite{krizhevsky2012-AlexNet}.
For example, deep learning significantly increases the number of layers in a neural network, which in turn multiplies the number of model parameters.
In general, larger models with more parameters can yield better performance than smaller networks \cite{coates2013-DL-COTS-HPC,dean2012-Large-scale-distributed,krizhevsky2012-AlexNet}.

Increasing the number of training examples and model parameters can improve the resulting accuracy significantly \cite{dean2012-Large-scale-distributed}
However, all those advances lead to substantially long training times of multiple days or even weeks \cite{jin2016-How-to-scale,chilimbi2014-Project-Adam}.
This makes experiments difficult and time-consuming \cite{yadan2013-Multi-GPU-Training,jiang2018-FiLayer} and is often a bottleneck to developing new, high accuracy model architectures \cite{iandola2016-Firecaffe}.

Consequently, parallelizing and distributing the training is necessary further to utilize all available computational power \cite{jiang2018-FiLayer,recht2011-Hogwild}.
Additionally, their sheer size prevents some datasets and models from fitting into the memory of a single machine \cite{recht2011-Hogwild,dean2012-Large-scale-distributed}.
This is especially prevalent when training on GPUs, as the available memory is much smaller than on CPUs \cite{dean2012-Large-scale-distributed}.
Circumventing memory bottlenecks can thus be an additional advantage when using distributed training.
% TODO: warum sind bisherige Ansätze (z.B. in PyTorch/Tensorflow) nicht gut genug?

% Hypthesis and overview over results/contributions
Different strategies for parallel and distributed training of neural networks have been developed.
Most techniques apply model parallelism, data parallelism, or hybrid approaches combining both approaches.
This paper discusses these different variants in detail, comparing their assets and drawbacks, and when to use which approach.
Additionally, we implement data parallel training using synchronous communication with all-reduce.
We test this training using convolutional neural networks on the well-known image classification problems MNIST and CIFAR10.
TODO: short summary of results/conclusion % TODO: something about the results from section 6/7

%% Organisation of this paper
This seminar paper is organized as follows:
The next section gives a short introduction to neural networks and their training.
After that, \autoref{sec:related_work} gives an overview of previous and related work on parallelizing the training of neural networks.
The different parallelization techniques are covered in detail in \autoref{sec:parallelizing_neural_networks}.
We discuss both model and data parallelism as well as hybrid approaches.
\autoref{sec:implementation} describes the implementation of data-parallel training using synchronous stochastic gradient descent (SGD).
The results are illustrated in \autoref{sec:experimental_results}.
Finally, \autoref{sec:conclusion} summarizes our contributions and results and gives an outlook on possible future work.

% section introduction (end)

\section{Preliminaries} % (fold)
\label{sec:preliminaries}
%TODO: look at preliminaries in \cite{paine2013-GPU-async-SGD} and \cite{chilimbi2014-Project-Adam}

% Deep Learning definition:
% \begin{itemize}
% 	\item use many layers
% 	\item learning hierarchical features from the raw input data instead of using hand-crafted features \cite{chilimbi2014-Project-Adam}
% \end{itemize}

This sections gives a short introduction to machine learning using convolutional neural networks, defining all necessary concepts used in the remainder of this paper.
More in-depth information can, for example, be found in \cite{goodfellow2016-DeepLearningBook}.

Neural networks consist of multiple neurons.
A neuron has some weighted connections, for example to previous neurons, with weights $w$ and a bias $b$.
It receives some input $x$ via its connections and computes a weighted sum $w^Tx+b$ which is then passed to an \emph{activation function} $\phi(\cdot)$.
Typical activation functions are $\tanh$, the logistic function, or more recently the rectified linear unit (ReLU).
The output $\phi(w^Tx+b)$ of the neuron can then be passed as input to the next neuron.

Multiple neurons can be organized into \emph{layers} where the output of each layer is passed as input to the next.
The input data -- for example the pixel values of a training image -- is passed to the first layer.
The network output is encoded in the last layer's output.
%
There are different types of layers depending on how the neurons are connected to the previous layer.
In computer vision and image classification, the most important types of layers are fully-connected layers, convolutional layers, and pooling-layers.

In \emph{fully-connected layers}, each neuron of the first layer is connected to each neuron in the second layer.
These layers consist of many connections which result in numerous parameters.
Fully-connected layers are typically used to determine the image class once the features have been extracted.

\emph{Convolutional layers} are based on the convolutional filtering used in digital image processing, for example in the Gaussian blur filter TODO cite? or edge detection filters like Sobel or Prewitt TODO cite?.
They use prior knowledge, for example the data hierarchy TODO cite, to reduce the number of parameters.
Convolutional layers consist of a small filter -- usually between 3x3 and 7x7 pixels -- that is shifted over the image.
At each pixel, the weighted sum of its surrounding pixels is computed and passed to the activation function.
Instead of connecting individual neurons with weighted connections, the filter weights are therefore reused for each pixel.
This reduces the number of parameters significantly, which in turn speeds up the training.
Convolutional layers are often used together with pooling layers as features extractors TODO cite.

\emph{Pooling layers} aggregate adjacent pixels to a single output pixel using either the sum, the average or the maximum of the pooled pixels.
Similarly to convolutional layers, the pooling operation is shifted over the image and apply repeatedly the the different parts of the image.
Pooling layers reduces the spatial size of the input image and generally improve the models robustness to noise and clutter in the input.
TODO look at Y-L. Boureau, J. Ponce, Y. LeCun. A Theoretical Analysis of Feature Pooling in Visual Recognition. ICML 2010

A \emph{loss function} is used to evaluate the accuracy by comparing the networks output to the actual, expected output.
Different loss functions for different types of problems exist.
For image classification, a typical loss function is the cross entropy loss.
%
In general, neural networks can be thought of as ``large parameterized function[s]'' \cite{paine2013-GPU-async-SGD} where all the weights and biases are model parameters that can be modified to improve the accuracy.
When training a neural network, we aim to minimize the loss by adjusting its parameters.

A widely used optimization strategy for training neural networks is \emph{stochastic gradient descent (SGD)}. % TODO cite?
Gradient descent is an optimization algorithm to find local minima by adjusting the parameters in the opposite direction of the gradient.
%
To compute the gradient of a neural network, we typically use the \emph{backpropagation} algorithm.
Backpropagation is based on the chain rule, a rule used to compute the derivative of function compositions.
%
Instead of computing the gradient over the entire training set at once, SGD partitions the shuffled training set into \emph{batches}.
For each batch, we compute an averaged gradient using back-propagation and update the model parameters accordingly.
%
A \emph{training epoch} consists of processing each batch in the training set once.
Usually, multiple epochs -- that is multiple passes through the training set -- are required to train a neural network to convergence \cite{chilimbi2014-Project-Adam}.

There are many different hyper-parameters to tune the neural network and its training.
In this paper, only the \emph{batch size}, that is the number of training instances per batch, is of importance.
A larger batch size generally reduces the training time per epoch while increasing the number of epochs required to reach a certain accuracy.
As we will see in \autoref{sec:parallelizing_neural_networks} and \ref{sec:experimental_results}, the batch size can also affect the scalability of some parallelization techniques.

% TODO position?
A dataset consists of two separate sets, one for training the network and one to test its accuracy an unseen inputs and to avoid overfitting.
Each of these sets contains pairs of an input instance and the correct label.
In this paper, we focus on the image classification problem, where given an image, the neural network shall correctly determine the class depicted in the image.

% TODO additional? GPUs for neural networks and why they work well, short MPI introduction

% section preliminaries (end)

\section{Related Work} % (fold)
\label{sec:related_work}

% section related_work (end)

\section{Parallelizing Neural Networks} % (fold)
\label{sec:parallelizing_neural_networks}

\subsection{Model Parallelism} % (fold)
\label{sub:model_parallelism}
% definition/how does it work
% pro/contra (compared to data-parallelism)

% subsection model_parallelism (end)

\subsection{Data Parallelism} % (fold)
\label{sub:data_parallelism}
% definition/how does it work in general
% different approaches: synchronous/asynchronous etc.
% pro/contra (compared to model-parallelism)

% subsection data_parallelism (end)

\subsection{Hybrid Approaches} % (fold)
\label{sub:hybrid_approaches}
% definition/how does it work
% motivation, pro/contra to just one parallelism technique

% subsection hybrid_approaches (end)

% section parallelizing_neural_networks (end)

% for implementation: why did I choose synchronous data parallelism with all-reduce? what's the advantage for parallel programming

\section{Implementation} % (fold)
\label{sec:implementation}
We implement and evaluate a synchronous, data-parallel training using MPI-All-Reduce to communicate and aggregate the gradients between different processes.
% TODO: reasons why this technique implemented and not e.g. model parallelism or asynch approaches.

We test our implementation with two CNNs:
\emph{LeNet5} \cite{lecun1998gradient} consists of three convolutional layers, two maximum pooling layers, and two fully-connected layers.
\emph{TorchNet}\footnote{\url{https://pytorch.org/tutorials/beginner/blitz/neural_networks_tutorial.html\#define-the-network}} consists of two convolutional layers, two maximum pooling layers, and three fully-connected layers.

These models are trained on the image classification datasets \emph{MNIST} \cite{lecun1998gradient} and \emph{CIFAR10} \cite{krizhevsky2009CIFAR10}.
MNIST contains 60,000 training images and 10,000 test images of handwritten digits.
CIFAR10 consists of 50,000 training images and 10,000 test images for ten different classes such as cat, airplane, or bird.

The training is implemented in Python v3.6.6 with PyTorch v1.1.0, Torchvision v0.3.0 and Numpy v1.16.4.
On the GPU, CUDA v10.0 is used for computation.
For communication, OpenMPI v3.1 with the python wrapper MPI4Py v3.0.2.
The datasets were converted to HDF5 format using HDF5 v1.10 (on ForHLR II) and v1.8.12 (on LSDF) and are read via H5Py v2.9.0.

We evaluated our implementation two different machines:
The \emph{LSDF} server consists of two Intel Xeon E5-2630 v3 processors running at 2.40 – 3.20~GHz with eight cores each and two threads per core.
Additionally, LSDF contains four NVIDIA Tesla K80 GPU, each offering 4992 NVIDIA CUDA cores and 24~GB of GDDR5 memory.
%
We, furthermore, used Intel Xeon E5-2660 v3 processors on the HPC-system \emph{ForHLR II}. % TODO dünne oder fette knoten?
Each processor runs at 2.60 – 3.30~GHz and consists of ten cores per processor and two threads per core
%
Both machines use Red Hat Enterprise Linux 7 as operating system.

% TODO: interesting/relevant implementation details?

% section implementation (end)

\section{Experimental Results} % (fold)
\label{sec:experimental_results}
% illustrate results with plots etc.
% describe and interpret results

% section experimental_results (end)

\section{Conclusion} % (fold)
\label{sec:conclusion}
% summarize contributions, reference to introduction/abstract
% summarize central results and implications
% revisit section key points
% Benefits and shortcomings of approach and results

% section conclusion (end)

% \begin{thebibliography}{1}
% % TODO
% \end{thebibliography}
\bibliographystyle{IEEEtran}
\bibliography{IEEEabrv,references}

\end{document}